---
title: 'Multivariate Analysis 01'
author: "劉佳欣 Rafe C. H. Liu"
output:
  html_document:
    css: styles.css
    fig_height: 7.5
    fig_width: 10
    theme: spacelab
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
---
```{r setup, include=FALSE, cache=FALSE}
library(knitr)      # dynamic report generation, RMarkdown
library(ggplot2)    # data Viz
library(reshape2)   # long and wide format
opts_chunk$set(echo = TRUE, comment="", message=FALSE, warning=FALSE)
```

[Contact](mailto:snexuz@gmail.com)｜[Linkedin](http://tw.linkedin.com/in/rafechliu)

# Data

```{r}
library(dplyr)

# new dataset
ubike <- read.csv(file = "./data/ubike-weather-big5.csv", 
                  fileEncoding = 'BIG-5',
                  stringsAsFactors = F)

ubike <- ubike %>% select(
  date, hour, sno, sarea, sna, lat, lng, tot, 
  sbi = avg.sbi, bemp = avg.bemp,
  temp, humidity, pressure
) %>% mutate(
  sbi = ceiling(sbi), bemp = ceiling(bemp),
  temp = temp %>% round(digits = 2), 
  humidity = humidity %>% round(digits = 2), 
  pressure = pressure %>% round(digits = 2)
)

# A galance of dataset
head(ubike)

tail(ubike)

# ETL: Extract-Transform-Load
str(ubike)

ubike <- ubike %>% 
  mutate(date = as.Date(date),
         sarea = as.factor(sarea))

```


# Descriptive Statistics
```{r}
summary(ubike)

ubike %>%
  group_by(sarea, sna) %>%
  summarise(mean(bemp), median(bemp), max(bemp), min(bemp), sd(bemp))
```

### Cross-Analysis

**Summary Statistics by Groups**

```{r}
library(dplyr)

ubike %>%
  filter(date == '2014-12-08', hour == '20') %>%
  group_by(sarea) %>%
  summarise(n())

## Data Transformation: Long-data format to Wide-data
library(reshape2)

ubike %>%
  filter(sna == '世貿二館' | sna == '信義廣場(台北101)') %>%
  # filter(sna %in% c('世貿二館', '信義廣場(台北101)')) %>%
  select(date, hour, sna, sbi) %>%
  dcast(formula = date + hour ~ sna, value.var = 'sbi') %>%
  head()
```

### Correlation

- http://www.statmethods.net/stats/correlations.html

```{r}
dat <- ubike %>%
  filter(sna == '世貿二館' | sna == '信義廣場(台北101)') %>%
  # filter(sna %in% c('世貿二館', '信義廣場(台北101)')) %>%
  select(date, hour, sna, sbi) %>%
  dcast(formula = date + hour ~ sna, value.var = 'sbi')

cov(dat[,3:4])
cor(dat[,3:4], method = 'pearson')
```

# EDA

**Exploratory Data Analysis**

```{r}
ubike %>%
  filter(sna == '世貿二館' | sna == '信義廣場(台北101)') %>%
  # filter(sna %in% c('世貿二館', '信義廣場(台北101)')) %>%
  select(date, hour, sarea, sna, sbi) %>%
  head(10)

## plots
library(ggplot2)

ubike %>%
  filter(sna == '世貿二館' | sna == '信義廣場(台北101)') %>%
  select(date, hour, sarea, sna, sbi) %>% 
  ggplot(aes(x = hour, y = sbi, group = sna, color = sna)) +
  geom_line() + facet_grid(date~.) + 
  theme(text = element_text(family = 'STHeiti')) # for MAC user
```

# Hypothesis Testing
### t-tests

- http://www.statmethods.net/stats/ttest.html

```{r}
# Dataset
dat <- ubike %>%
  filter(sna == '世貿二館' | sna == '信義廣場(台北101)') %>%
  # filter(sna %in% c('世貿二館', '信義廣場(台北101)')) %>%
  select(date, hour, sna, sbi) %>%
  dcast(formula = date + hour ~ sna, value.var = 'sbi')
```

檢定兩區域平均租借數是否相等 (Independent 2-group sample t-test)

- H0: u1 = u2 vs H1: u1 != u2 
- p is small, reject H0, u1 != u2

```{r}
t.test(dat$世貿二館, dat$`信義廣場(台北101)`) 

## results are stored in the list
rt <- t.test(dat$世貿二館, dat$`信義廣場(台北101)`) 
rt$p.value
```


### ANOVA

- http://www.statmethods.net/stats/anova.html
- http://programmermagazine.github.io/201311/htm/article3.html

```{r}
# Dataset
dat <- ubike %>%
  filter(sna %in% c('世貿二館', '世貿三館', '信義廣場(台北101)', '市民廣場')) %>%
  select(date, hour, sna, sbi) 

head(dat)
summary(dat)
```

<br>
**Q: 探討四車站(sna)的平均租借數(sbi)租借數趨勢**
<br>

```{r}
# EDA
dat %>% 
  ggplot(aes(x = hour, y = sbi, group = sna, colour = sna)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) + theme(text = element_text(family = 'STHeiti')) 

dat %>% group_by(hour, sna) %>%
  summarise(mean(sbi)) %>%
  ggplot(aes(x = hour, y = `mean(sbi)`, group = sna, colour = sna)) +
  geom_line() + theme(text = element_text(family = 'STHeiti')) 

## 檢定四車站(sna)的平均租借數(sbi)是否相等
dat %>%
  ggplot(aes(x = sna, y = sbi)) +
  geom_boxplot() + theme(text = element_text(family = 'STHeiti')) 

## Ho: u1 = u2 = u3 = u4, H1: 全不相等
## results
fit <- aov(sbi ~ sna, data = dat) 
fit

## summary table: 
# WARNING: R provides Type I sequential SS, 
# not the default Type III marginal SS reported by SAS and SPSS. 
# Rej Ho, 
summary(fit)

par(mfrow=c(2,2))
plot(fit)

# post-hoc test: Multiple Comparisons
TukeyHSD(fit)
```

### MANOVA

- https://rpubs.com/aaronsc32/manova-test-statistics

**Q: 同時對四車站(sna)的平均租借數(sbi)及空位數(bemp)進行檢定**

```{r}
dat <- ubike %>%
  filter(sna %in% c('世貿二館', '世貿三館', '信義廣場(台北101)', '市民廣場')) %>%
  select(date, hour, sna, sbi, bemp) 

head(dat)
summary(dat)

# manova
fit <- manova(cbind(sbi,bemp) ~ sna, data = dat) 

summary(fit) 

# 等同做兩次 anova
summary.aov(fit) # univariate ANOVA tables
summary(fit, test = "Wilks") # ANOVA table of Wilks' lambda
```

# Linear Models

- http://www.statmethods.net/stats/regression.html

**Q: 平均租借數(sbi)影響因素**

```{r}
dat <- ubike %>% filter(sna == '世貿二館') 

head(dat)
summary(dat)

## correlation
cor(dat %>% select(sbi, temp, humidity, pressure))

# https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html
# install.packages('corrplot')
library(corrplot)
cor(dat %>% select(sbi, temp, humidity, pressure)) %>% corrplot()

# Model
fit <- lm(sbi ~ humidity, data = dat)
summary(fit)

# Other useful functions 
coefficients(fit) # model coefficients
confint(fit, level=0.95) # CIs for model parameters 
fitted(fit) # predicted values
residuals(fit) # residuals
anova(fit) # anova table 
vcov(fit) # covariance matrix for model parameters 
influence(fit) # regression diagnostics

## compare with ano function
aov(sbi ~ humidity, data = dat) %>% summary()

## diagnostic plots 
par(mfrow=c(2,2))
plot(fit)

## discussions
lm(temp ~ pressure, data = dat) %>% summary()
lm(temp ~ pressure + humidity, data = dat) %>% summary()

lm(sbi ~ temp + pressure + humidity, data = dat) %>% summary()
```

# Generalized Linear Models 

廣義線性模型是傳統線性模型的延伸，可使用非常態分布的資料(例如視為離散的計數或測量比例) 建立模型。 

- `?glm`
- http://www.theanalysisfactor.com/r-tutorial-glm1/
- http://stats.idre.ucla.edu/r/
- http://www.r-tutor.com/elementary-statistics/logistic-regression/estimated-logistic-regression-equation


```{r}
# 製作有車沒車機率資料集
ubike %>% select(sarea, sna, tot, bemp) %>%
  mutate(hv_cars = round(bemp/tot, digits = 4)) %>%
  ggplot(aes(x = hv_cars)) + geom_histogram(binwidth = 0.05) 

# 有車，空車率 >60%，有車
dat <- ubike %>% 
       mutate(hv_cars = round(bemp/tot, digits = 4),
              hv_cars_yn = ifelse(hv_cars > 0.6, 1, 0))

dat <- dat %>% filter(sarea == '信義區')

head(dat)

summary(dat)

### 交叉分析

library(reshape2)

dcast(formula = sna ~ hv_cars_yn, dat)

dcast(formula = sna ~ hv_cars_yn, dat, 
      fun.aggregate = function(x) round(mean(x), digit = 4), 
      value.var = 'hv_cars')
```

### logistic regression

- Generalized Linear Models - binomial family

```{r}
logitreg <- glm(hv_cars_yn ~ temp + humidity + pressure, data = dat, family = "binomial")
summary(logitreg)

## CIs using profiled log-likelihood
confint(logitreg)

## CIs using standard errors
confint.default(logitreg)

## odds ratios 
exp(coef(logitreg))

## Likelihood
logLik(logitreg)
```

<div style="text-align: center;">
<br> Thanks!<br>
Rafe C.H. Liu｜[Contact](mailto:snexuz@gmail.com)｜[Linkedin](http://tw.linkedin.com/in/rafechliu)
</div>