---
title: "R - Homework assignments example"
author: "Rafe C. H. Liu"
date: "2016/10/17"
output:
  html_document:
    css: styles.css
    fig_height: 7.5
    fig_width: 10
    theme: spacelab
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, cache=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(plotly)

Sys.setlocale("LC_TIME", "C")
```


<hr>

# Data project
  1. 問題定義 (Problem Statement)
  2. 專案資料清理 (Data Wrangling)
  3. 探索性資料分析 (Exploratory Data Analysis, EDA) => Dashboard
  4. 驗證性因素分析 (Confirmatory data analysis) => Modeling
    - Forecasting: 
      - Regression Models
      - Time Series Models, Moving Average...
      - Machine Learning / Data Mining
  
<hr>
  
# Expected Outcomes and Project Demo

## Data
### Data wrangling

```{r comment="", warning=F}
POS <- read.csv("POS.csv", fileEncoding = "big5", stringsAsFactors = F)
POS <- POS[-c(8,9)]
colnames(POS) <- c("date", "shift", "total", "mis", "sub", "orders", "cups")

POS <- POS %>% mutate( 
                date = as.Date(date),
                total = gsub(",", replacement = "", total) %>% as.numeric(), # replace "," with ""
                mis = as.numeric(mis),
                sub = gsub(",", replacement = "", sub) %>% as.numeric(),
                orders = as.numeric(orders),
                cups = as.numeric(cups)) 

# replace NA with 0
POS[is.na(POS)] <- 0

head(POS) %>% knitr::kable()
```

### Generate new vars
```{r comment="", warning=F}
POS <- POS %>% 
   mutate(#shift = ifelse(shift == "早", 1, 2) %>% as.factor(),
          shift = factor(shift, levels = c("早", "晚")),
          year = format(date, format = "%Y"),
          month = format(date, format = "%m"),
          day = format(date, format = "%d"),
          weekdays = format(date, format = "%a"),
          cash = total + mis - sub) 

head(POS) %>% knitr::kable()
```

## Explantory data analysis

### Desciptive analysis
```{r comment="", warning=F}
summary(POS)

```

### Cross Analysis
```{r comment="", warning=F}
POS %>% group_by(month, weekdays) %>% summarise(cash.total = sum(cash)) %>% head(10) %>% knitr::kable()

POS %>% dcast(month + day + weekdays ~ shift, fun.aggregate = sum, value.var = "cash", margins = "shift") %>%
  head(10) %>% knitr::kable()

```


### Some graphics
```{r comment="", warning=F}

POS %>%
  group_by(shift) %>%
    summarise(shiftsales = sum(total),
              percentage = shiftsales / sum(547987+401899)) %>%
  ggplot(data =., aes(x = shift, y = percentage, fill = percentage)) + 
  geom_bar(stat = "identity", width = 1) 

POS %>%
  group_by(shift) %>%
    summarise(shiftsales = sum(total),
              percentage = shiftsales / sum(547987+401899)) %>%
  ggplot(data =., aes(x = "", y = percentage, fill = percentage)) + 
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y")

POS %>%
group_by(weekdays) %>%
    summarise( weekdaysales = sum(total)) %>%
    ggplot(data = ., aes(x = weekdays, y = weekdaysales )) +
    geom_bar(stat="identity" )


POS %>% 
  group_by(date) %>% summarise(cash.total = sum(cash)) %>%
  ggplot(aes(x = date, y = cash.total)) +
  geom_line()


POS %>% 
  group_by(month, day) %>% summarise(cash.total = sum(cash)) %>%
  ggplot(aes(x = day, y = cash.total, group = month, color = month)) +
  geom_line()


# plotly demo
p <- POS %>% 
  group_by(month, day) %>% summarise(cash.total = sum(cash)) %>%
  ggplot(aes(x = day, y = cash.total, group = month, color = month, fill = month)) +
  geom_bar(stat = "identity") + 
  facet_wrap(~ month) 

ggplotly(p)


# plotly dropdown

POS %>% dcast(day ~ month, fun.aggregate = sum, value.var = "cash") %>%
  plot_ly(x = ~day, type = 'bar') %>%
  add_trace(y = ~`06`, name = "Jun.") %>%
  add_trace(y = ~`07`, name = "Jul.", visible = F) %>%
  add_trace(y = ~`08`, name = "Aug.", visible = F) %>%
  add_trace(y = ~`09`, name = "Sep.", visible = F) %>%
  layout(
  title = "Drop down menus - Styling",
  updatemenus = list(
    list(
      x =0.9, y = 0.75, # dropdown mean 的位置
      buttons = list(
        list(method = "restyle",
             args = list("visible", list(TRUE, FALSE, FALSE, FALSE)),
             label = "June."),
 
        list(method = "restyle",
             args = list("visible", list(FALSE, TRUE, FALSE, FALSE)),
             label = "July."),
    
        list(method = "restyle",
             args = list("visible", list(FALSE, FALSE, TRUE, FALSE)),
             label = "August."),

        list(method = "restyle",
             args = list("visible", list(FALSE, FALSE, FALSE, TRUE)),
             label = "September.")
        ))
  ))

```

## Modeling
### Moving Average Forecasting
```{r comment="", warning=F}
library(forecast)

dat <- POS %>% group_by(date) %>% summarise(cash.total = sum(total)) 
nrow(dat)
# time series obj.
tdat <- ts(dat$cash.total[11:122], frequency = 7) # from mon to sun
tdat %>% print(calendar = TRUE) 

library(forecast)

moving_average <- ma(tdat[1:70], order=3) %>% forecast(h = 7)
moving_average_accuracy <- accuracy(moving_average, tdat[71:78])
moving_average; moving_average_accuracy

plot(moving_average, ylim = c(6000, 15000), lwd = 2, col = "red")
lines(tdat[1:78], lwd = 1)

```

# CONTACT

- Mobile: **0912-120811**
- snexuz@gmail.com
- [Linkedin](http://tw.linkedin.com/in/rafechliu)

